---
title: "Practical Machine Learning Project"
output: html_document
author: Leon Ma
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

##Data

The training data for this project are available at:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available at:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

###Preprocessing

```{r Preprocessing}

library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(e1071)
library(randomForest)
set.seed(1)

train.url <-
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test.url <- 
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

path <- paste(getwd(),"/", "machine", sep="")
train.file <- file.path(path, "machine-train-data.csv")
test.file <- file.path(path, "machine-test-data.csv")

```
The files are read into memory. Various indicators of missing data (i.e., "NA", "#DIV/0!" and "") are all set to NA so they can be processed.

```{r download, echo=FALSE}

if (!file.exists(train.file)) {
        download.file(train.url, destfile=train.file)
}
if (!file.exists(test.file)) {
        download.file(test.url, destfile=test.file)
}

trainingset  <- read.csv(train.file, na.strings=c("NA","#DIV/0!",""))
testingset  <- read.csv(test.file, na.strings=c("NA","#DIV/0!",""))

```

###Remove unecessary columns

```{r remove, echo=TRUE}
# Drop the first 7 columns as they're unnecessary for predicting.
train.data.clean1 <- trainingset[,8:length(colnames(trainingset))]
test.data.clean1 <- testingset[,8:length(colnames(testingset))]

# Drop colums with NAs
train.data.clean1 <- train.data.clean1[, colSums(is.na(train.data.clean1)) == 0] 
test.data.clean1 <- test.data.clean1[, colSums(is.na(test.data.clean1)) == 0] 

# Check for near zero variance predictors and drop them if necessary
nzv <- nearZeroVar(train.data.clean1,saveMetrics=TRUE)
zero.var.ind <- sum(nzv$nzv)

if ((zero.var.ind>0)) {
        train.data.clean1 <- train.data.clean1[,nzv$nzv==FALSE]
}
```
###Cross Validation

The training data is divided into two sets. This first is a training set with 70% of the data which is used to train the model. The second is a validation set used to assess model performance.

```{r crossValidation, echo=TRUE}

in.training <- createDataPartition(train.data.clean1$classe, p=0.70, list=F)
train.data.final <- train.data.clean1[in.training, ]
validate.data.final <- train.data.clean1[-in.training, ]

```

##MODEL

The model fit using the training data is tested against the validation data. Predicted values for the validation data are then compared to the actual values. This allows forecasting the accuracy and overall out-of-sample error, which indicate how well the model will perform with other data.

```{r model2, echo=TRUE}

model2 <- randomForest(classe ~. , data=train.data.final, method="class")

# Predicting:
prediction2 <- predict(model2, validate.data.final, type = "class")

# Test results on subTesting data set:
confusionMatrix(validate.data.final$classe, prediction2)

```


##Result

```{r result, echo=TRUE}

predictfinal <- predict(model2, testingset, type="class")
predictfinal

```

##Apendix

```{r appendix, echo=TRUE}
model1 <- rpart(classe ~ ., data=train.data.final, method="class")

# Predicting:
prediction1 <- predict(model1, validate.data.final, type = "class")

# Plot of the Decision Tree
rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)

confusionMatrix(prediction1, validate.data.final$classe)
```